% http://people.su.se/~ma/R_intro/R_intro.pdf

Time-series

### Data in rows
Time-series data often appear in a form where series are in rows. As an example
I use data from the Swedish Consumer Surveys by the National Institute of
Economic Research containing three series: consumer condence index, a macro
index and a micro index.

First I saved the original data in a file in text-format. Before using the data as
input in R I used a text editor and kept only the values for the rst three series
separated with spaces. The series are in rows. The values of the three series are
listed in separate rows without variable names.

To read this file <macro.txt>, the following code puts the data in a matrix of
3 columns and 129 rows with help of <scan> and <matrix> before dening a
time-series object starting in 1993 with frequency 12 (monthly). The series are
named as <cci>,<macro.index> and <micro.index>. Notice that <matrix>
by default lls in data by columns.

\begin{verbatim}
> FILE <- "http://people.su.se/~ma/R_intro/macro.txt"
> macro <- ts(matrix(scan(FILE), 129, 3), start = 1993,
+ frequency = 12, names = c("cci", "macro.index",
+ "micro.index"))
\end{verbatim}
Here I give an example for creating lag values of a variable and adding it to a
time-series data set. See also <diff> for computing dierences.
Let us create a new time series data set, with the series in the data frame
<macro> adding lagged <cci> (lagged by 1 month). The function <ts.union>
puts together the series keeping all observations while <ts.intersect> would
keep only the overlapping part of the series.
> macro2 <- ts.union(macro, l.cci = lag(macro[,
+ 1], -1))
You can use the function aggregate to change the frequency of you time-series
data. The following example converts the frequency data. nfrequency=1 yields
annual data. FUN=mean computes the average of the variables over time. The
default is <sum>.
> aggregate(macro, nfrequency = 1, FUN = mean)
Time Series:
Start = 1993
End = 2002
Frequency = 1
cci macro.index micro.index
1993 -19.7583333 -33.3833333 -13.241667
1994 -0.2916667 14.1666667 -5.158333
